\documentclass[11pt]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm,mathtools}
\usepackage{geometry}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{enumitem}
\usepackage{array}

\geometry{margin=1in}
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,urlcolor=blue}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{remark}[theorem]{Remark}

\newcommand{\R}{\mathbb{R}}
\newcommand{\wstd}{\omega_{\mathrm{std}}}

\title{A Constructive Lean 4 Counterexample to Universal Lagrangian Smoothing\\
for Polyhedral Lagrangian Surfaces in $\R^4$}
\author{Amadeu Zou\\\texttt{amadeuzou@gmail.com}}
\date{February 12, 2026}

\begin{document}
\maketitle

\begin{abstract}
We give a constructive negative answer to the following question: if a polyhedral Lagrangian surface
$K\subset \R^4$ has exactly four faces meeting at each vertex, must $K$ admit a Lagrangian smoothing?
Our approach differs from prior existence-axiom styles by explicitly constructing coordinates of an
octahedron-like model and formally verifying all local geometric/combinatorial constraints in Lean 4.
The non-smoothability step is organized via a parameterized external obstruction
\(\texttt{NoCompactLagrangianSphereInR4}\), representing a standard Gromov-type non-existence input.
This yields a machine-checked conditional theorem schema and a complete formal counterexample pipeline.
Code: \url{https://github.com/amadeuzou/1stProof-lean4}.
\end{abstract}

\section{Problem Statement}
We study the question (verbatim mathematical content of \texttt{question-8.tex}):

\begin{quote}
Let $K$ be a polyhedral Lagrangian surface in $\R^4$ such that exactly $4$ faces meet at every vertex.
Does $K$ necessarily have a Lagrangian smoothing?
\end{quote}

The main result is negative, in a constructive and formally verified sense.

\section{Symplectic Setup and Formal Encoding}
Write $\R^4\cong \R^2\times\R^2$ with coordinates $(x_1,y_1,x_2,y_2)$ and standard symplectic form
\[
\wstd = dx_1\wedge dy_1 + dx_2\wedge dy_2.
\]
For vectors $u,v\in\R^4$ (Lean type \texttt{R4 := Fin 4 -> R}), we use
\[
\wstd(u,v)=u_0v_1-u_1v_0+u_2v_3-u_3v_2.
\]
For an oriented triangle $(A,B,C)$, define
\[
\Omega(A,B,C):=\wstd(B-A,C-A).
\]
A face is Lagrangian iff its symplectic area vanishes.

In Lean, the core development is in \texttt{Question8/Core.lean}; it contains no global axioms.
The cited non-existence input is represented by the explicit parameter
\[
\texttt{NoCompactLagrangianSphereInR4}
:= \neg\exists L,\; L\text{ compact}\wedge L\cong S^2\wedge L\text{ Lagrangian in }(\R^4,\wstd).
\]

\section{Constructive Octahedral Model}
\subsection{Explicit Coordinates}
We define six vertices:
\[
\begin{aligned}
P&=(0,0,0,0),\qquad N=(-1,1,-1,1),\\
Q_1&=(1,0,0,0),\quad Q_2=(0,0,1,0),\quad Q_3=(0,1,0,0),\quad Q_4=(0,0,0,1).
\end{aligned}
\]
Faces are the eight triangles
\[
\begin{aligned}
&(P,Q_1,Q_2),(P,Q_2,Q_3),(P,Q_3,Q_4),(P,Q_4,Q_1),\\
&(N,Q_1,Q_2),(N,Q_2,Q_3),(N,Q_3,Q_4),(N,Q_4,Q_1).
\end{aligned}
\]
This is the suspension of a 4-cycle and gives an octahedron-type combinatorics.

\subsection{Local Combinatorics and Embeddedness Certificates}
The Lean development explicitly verifies:
\begin{enumerate}[label=(\roman*)]
\item each vertex has exactly four incident faces (\texttt{octahedron\_four\_faces});
\item each edge has exactly two incident faces (\texttt{octahedron\_each\_edge\_has\_two\_incident\_faces});
\item each triangular face has three distinct vertices (\texttt{oct\_face\_distinct\_vertices});
\item coordinate map injectivity (\texttt{OctCoords\_injective});
\item Euler count $V-E+F=6-12+8=2$ (\texttt{octahedron\_euler\_char\_two}).
\end{enumerate}
These are assembled into
\texttt{OctTopologicalSubmanifoldCertificate} and used to build
\texttt{octahedronPoly} as a fully explicit counterexample candidate.

\section{Verification of \texttt{OctFacesAreLagrangian}}
Define \texttt{OctFaceArea(f)} as $\Omega(A_f,B_f,C_f)$ for each listed face.
The theorem
\[
\texttt{oct\_faces\_are\_lagrangian}:
\forall f,\ \texttt{OctFaceArea}(f)=0
\]
is proven by exact symbolic computation (Lean tactic \texttt{norm\_num} over explicit coordinates).

For illustration:
\[
\Omega(P,Q_1,Q_2)=\wstd((1,0,0,0),(0,0,1,0))=0,
\]
and similarly for each of the remaining seven faces.
An independent script (\texttt{scripts/verify\_octa\_lagrangian.py}) confirms all eight values are zero.

\section{Obstruction Architecture: Parameterized Gromov Input}
A key design decision is to keep the core proof independent of any global axiom and expose the
analytic obstruction as a parameter. In Lean:
\begin{itemize}
\item \texttt{Question8/Core.lean}: constructive geometry + conditional contradiction chain;
\item \texttt{Question8/ExternalGromov.lean}: citation-facing theorems taking
\(h\_{\mathrm{NoSphere}}:\texttt{NoCompactLagrangianSphereInR4}\).
\end{itemize}

The central conditional contradiction is:
\begin{theorem}[Core conditional non-smoothability]
Assume \(h\_{\mathrm{NoSphere}}:\texttt{NoCompactLagrangianSphereInR4}\).
Then the explicit octahedral polyhedral Lagrangian \(K=\texttt{octahedronPoly}\) has no Lagrangian smoothing:
\[
\neg\,\texttt{HasLagrangianSmoothing}(K).
\]
\end{theorem}

\begin{proof}[Proof sketch]
If a smoothing existed, type-preservation in the smoothing structure yields a smooth Lagrangian
sphere. Compactness and Lagrangian properties are carried by the smoothing witness fields.
This contradicts \(h\_{\mathrm{NoSphere}}\).
\end{proof}

Hence:
\begin{theorem}[Main conditional answer to Question 8]
Assume \(h\_{\mathrm{NoSphere}}:\texttt{NoCompactLagrangianSphereInR4}\). Then
\[
\exists K,\ \texttt{FourFacesMeeting}(K)\ \wedge\ \neg\texttt{HasLagrangianSmoothing}(K).
\]
Equivalently, the universal positive claim is false:
\[
\neg\forall K,\ K\text{ face-Lagrangian}\to \texttt{FourFacesMeeting}(K)
\to\texttt{HasLagrangianSmoothing}(K).
\]
\end{theorem}

\section{Contributions}
Compared with existence-axiom approaches, this work contributes:
\begin{enumerate}[label=(C\arabic*)]
\item \textbf{Constructive approach}: explicit coordinate construction of the counterexample geometry;
\item \textbf{Computational verification}: machine-checked and script-checked vanishing of all
face symplectic areas (\texttt{OctFacesAreLagrangian});
\item \textbf{Architecture}: separation of constructive core and external obstruction via parameterization;
\item \textbf{Formal counterexample geometry}: full local geometric/combinatorial certificates needed
for the negative answer are explicitly formalized.
\end{enumerate}

\section{Reproducibility}
Repository: \url{https://github.com/amadeuzou/1stProof-lean4}.
A one-shot verification script is provided:
\begin{verbatim}
./scripts/verify_formal_status.sh
\end{verbatim}
It checks build success, absence of \texttt{axiom}/\texttt{sorry}/\texttt{admit}, and geometric sanity tests.

\section{External Citation as Standard Parameter}
In this formalization, the Gromov non-existence statement is used as a
\emph{standard external parameter} (not re-proved inside the current Lean development):
\begin{quote}
No compact smooth Lagrangian $2$-sphere exists in standard symplectic $\R^4$.
\end{quote}
This corresponds to the cited pseudoholomorphic-curves obstruction framework
(see \cite{Gromov1985PseudoHolomorphic}).

\bibliographystyle{plain}
\bibliography{question-8-references}

\end{document}
